### 2021-06-15 paper/NIHGrant.tex
Refactor script and add clarification edits to Aim #3

The \textbf{objective} of this proposal is to directly and more completely characterize the perceptual experience of tinnitus using the novel application of a primary method for uncovering latent perceptual representations: reverse correlation.
Tinnitus is the perception of ringing, buzzing, or hissing in the ears, which can range from mild and transient to debilitating.
In the U.S. alone, it affects 45 million people per year, 30\% of which report their tinnitus to be moderate to severe.
A primary form of treatment habituation therapy that retrains a patient’s nervous system by habituating it to external sounds that mimic the perceptual experience of the tinnitus sound.
Habituation therapies require accurate estimates of patients’ tinnitus experience in order to promote patient outcomes.
There is a pressing need for new approaches to uncovering individuals’ perceptual experience of tinnitus,
because current estimation methods rely on strongly reductionist assumptions concerning the nature of tinnitus sounds, 
and produce results that are correspondingly biased and incomplete.
Reverse correlation holds promise in this regard,
as a primary method used to characterize neural tuning that is known to allow for unconstrained and unbiased estimation of latent neural representations
(\eg neural receptive fields) \cite{ringachReverseCorrelationNeurophysiology2004, nishimotoReceptiveFieldProperties2006},
as well as higher-level cognitive representations, including psychophysical kernels that drive the top-down processes of perception 
(\eg face or phoneme recognition) \cite{gosselinSuperstitiousPerceptionsReveal2003,smithMeasuringInternalRepresentations2012,neriReceptivePerceptiveFields2006,ahumadaStimulusFeaturesSignal1971},
and even abstract psychological categories (\eg “male” vs. “female” faces)
\cite{brinkmanVisualisingMentalRepresentations2017,manginiMakingIneffableExplicit2004,ponsotCrackingSocialCode2018},
all on the basis of fairly simple stimulus-response data \cite{ljungMeasureLackFit1978,marmarelisWhiteNoiseMethodSystem1978}.
Despite its wide acceptance in other areas of inquiry, reverse correlation has not been brought to bear with respect to uncovering tinnitus experience.

Our \textbf{long-term goal} is to directly uncover and more completely characterize the perceptual experience of tinnitus using the demonstrated capabilities of the reverse correlation approach.
Our guiding hypothesis is that reverse correlation will produce estimates of tinnitus percepts that are substantially more rich, individualized and unbiased than currently possible using existing approaches. The rationale is that high quality characterization of tinnitus percepts can promote detailed assessment of tinnitus, helping to explain individual variability in tinnitus experience, and ultimately informing individualized treatment through enhanced habituation therapies. The proposed work will simultaneously improve the efficiency of the reverse correlation method itself, further enhancing its potential for adoption in clinical settings. We will test our central hypothesis and attain our objective via the following specific aims:

\begin{description}
	\item[Aim \#1: Uncover and reconstruct cognitive representations of tinnitus]{} 
\end{description}

To reconstruct perceptual representations of tinnitus using a novel reverse correlation procedure. Participants with tinnitus will be presented with richly varying stimuli in the context of a yes-no behavioral response paradigm. Perceptual representations can be estimated by regressing observed responses against the stimuli over many trials. Reconstructed representations will be compared to conventional estimates of tinnitus from the same participants. Expected outcome: reconstructed representations of tinnitus will more detailed and less biased than any those obtained through conventional methods.

\begin{description}
	\item[Aim \#2: Develop efficient reconstruction algorithm for cognitive representations using compressive sensing]{} 
\end{description}

To enhance the efficiency of the reverse correlation procedure using compressive sensing, which substantially reduces the number of trials required to accurately reconstruct perceptual representations. We will use compressive sensing to more efficiently and more accurately reconstruct cognitive representations using data obtained in Aim \#1. Expected outcome: compressive sensing is expected to improve efficiency of the reverse correlation procedure by up to 90\%.
% The reverse correlation and compressive sensing framework will be released
% as an open-source software implementation for use by the research community.
% Our pilot data demonstrates that compressive sensing affords a 90\% decrease
% in the number of trials for equivalent resultant latent representation.

\begin{description}
	\item[Aim \#3: Characterize and interpret cognitive representations of tinnitus]{} 
\end{description}

To characterize variability in perceptual representations of tinnitus between participants and over time. We will use unsupervised machine learning techniques to reveal latent similarities between tinnitus representations, and subsequently cluster those into distinct categories. Expected outcome: Unsupervised machine learning will identify distinct categories of tinnitus, which will allow us to further improve the effiency and accuracy of uncovering tinnitus representations through improved stimulus generation and representation classification.

% Dimensionality reduction algorithms will reveal latent similarities between
% tinnitus representations which will be subtyped using a data-driven approach.

% Our compressive sensing framework will allow for rapid data acquisition from participants.
% We will characterize cognitive representations of tinnitus between participants and over time.
% What is the space of tinnitus cognitive representations?
% Are there subtypes or clusters on a lower dimensional manifold?
% We will employ dimensionality reduction and clustering algorithms to represent and visualize the data,
% providing direct insight into commonalities and variability of tinnitus experience in and between subjects.

@@

\subsection*{C. Aim \#1: Uncover and reconstruct cognitive representations of tinnitus}

@@

\subsubsection*{C.1. Method}

@@

Surveys of tinnitus patients have identified sixteen qualitative subtypes of tinnitus
\cite{vajsakovicPrinciplesMethodsPsychoacoustic2021,meikleTinnitusArchiveArchive2004,stoufferCharacterizationTinnitusTinnitus1990},
but these categories are limited by the fuzziness of linguistic categories and the musical ability of the tinnitus patients.
Tinnitus subtypes are used to better explain and quantify the condition,
however qualitative subtypes are ill-defined.
While one tinnitus patient we interviewed, who has a degree in music performance,
was able to identify her tinnitus as sounding like a particular note played on the crotales,
a niche percussion instrument,
most tinnitus patients lack the depth of vocabulary to explain their tinnitus percept that precisely.
Accordingly, linguistic description of tinnitus can be vague and overlapping.
For instance, two of the subtypes in \cite{vajsakovicPrinciplesMethodsPsychoacoustic2021} are ``high tension wire'' and ``transformer noise.''
Is ``high tension wire'' meant in the colloquial sense of high-voltage power lines,
or does it mean the reverberation of a tightly-pulled string?
If the latter, then how is the sound of a high-voltage power line
disambiguated from the sound of a distribution transformer,
which is
mounted on a utility pole and connected directly to the high-voltage wires?
Instead of linguistic categories, we propose to use latent variable analysis
to reveal data-driven subtypes of tinnitus.

@@

\begin{figure}[h] % Figure at bottom of the page ([b] argument, could be "t" for top or "h" for here)
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/variational-autoencoder.png}
	\caption{Diagram of a variational autoencoder (VAE) learning a representation of the digit ``2''.
	VAEs assume that the data, $x$, are generated by a directed graphical model, $p_\theta(x|z)$,
	and that the encoder is learning an approximation $q_\phi(z|x)$, where $\phi$ and $\theta$
	denote the parameters of encoder and decoder respectively.
	The latent representation, $z$, is interpreted as originating from a multivariate Gaussian distribution.
	By sampling from the latent distribution, new data can be generated. Figure from \cite{IntroductionAutoencoders2020}.}
	\label{fig:vae}
\end{figure}

@@

While we anticipate remarkable performance gains in the number of samples required for accurate representation
through compressive sensing,
it may be clinically useful to identify only the subtype of tinnitus
(\ie which cluster on the lower-dimensional manifold the representation belongs to)
to some convergent accuracy.
If only the subtype is needed,
we believe the number of samples required for convergent accuracy can be further reduced.
We will devise an algorithm to perform the data-driven tinnitus subtyping ``online''
while the subject performs the task.
When convergent accuracy is reached, the task will automatically conclude.
By this process, both patient and clinician time is optimized, while still providing high-fidelity quantitative medical data
for use in diagnosis and treatment.

@@

>>>>

The \textbf{objective} of this proposal is to directly and more completely characterize the perceptual experience of tinnitus using the novel application of a primary method for uncovering latent perceptual representations: reverse correlation. Tinnitus is the perception of ringing, buzzing, or hissing in the ears, which can range from mild and transient to debilitating. In the U.S. alone, it affects 45 million people per year, 30\% of which report their tinnitus to be moderate to severe. A primary form of treatment is habituation therapy, that retrains a patient's nervous system by habituating its response to sounds that mimic the perceptual experience of the tinnitus sound. Habituation therapies require accurate estimates of patients' tinnitus experience in order to promote patient outcomes. There is a pressing need for new approaches to uncovering individuals' perceptual experience of tinnitus, because current estimation methods rely on strongly reductionist assumptions concerning the nature of tinnitus sounds, and produce results that are correspondingly biased and incomplete. Reverse correlation holds promise in this regard, as a primary method used to characterize neural tuning that is known to allow for unconstrained and unbiased estimation of latent neural representations (e.g., neural receptive fields) [29, 33], as well as higher-level cognitive representations, including psychophysical kernels that drive the top-down processes of perception (e.g., face or phoneme recognition) [1, 10, 28, 34], and even abstract psychological categories (e.g., male vs. female faces) [6, 22, 32], all on the basis of fairly simple stimulus-response data [20, 23]. Despite its wide acceptance in characterizing neural and cognitive representations, reverse correlation has not been brought to bear with respect to uncovering tinnitus experience.

Our long-term goal is to directly uncover and more completely characterize the perceptual experience of tinnitus using the demonstrated capabilities of the reverse correlation approach. Our guiding hypothesis is that reverse correlation will produce estimates of tinnitus percepts that are substantially more rich, individualized and unbiased than those currently possible using existing approaches. The rationale is that high quality characterization of tinnitus percepts can promote detailed assessment of tinnitus, helping to explain individual variability in tinnitus experience, and ultimately informing individualized treatment through enhanced habituation therapies. The proposed work will furthermore improve the efficiency of the reverse correlation method itself, further enhancing its potential for adoption in clinical settings. We will test our guiding hypothesis and attain our objective via the following specific aims:

\subsection*{Aim \#1: Uncover and reconstruct cognitive representations of tinnitus}

To reconstruct perceptual representations of tinnitus using a novel reverse correlation procedure. Participants with tinnitus will be presented with richly varying stimuli in the context of a yes-no behavioral response paradigm. Perceptual representations can be estimated by regressing observed responses against the stimuli over many trials. Reconstructed representations will be compared to conventional estimates of tinnitus from the same participants. 
Expected outcome: reconstructed representations of tinnitus will be more detailed and less biased than any those obtained through conventional methods.
\subsection*{Aim \#2: Develop efficient reconstruction algorithm for cognitive representations using compressive sensing}

To enhance the efficiency of the reverse correlation procedure using compressive sensing, which substantially reduces the number of trials required to accurately reconstruct perceptual representations. We will use compressive sensing to more efficiently and more accurately reconstruct cognitive representations using data obtained in Aim \#1. Expected outcome: compressive sensing is expected to improve efficiency of the reverse correlation procedure by up to 90\%.

\subsection*{Aim \#3: Characterize and categorize cognitive representations of tinnitus}

To characterize variability in perceptual representations of tinnitus between participants and over time. We will use unsupervised machine learning techniques to reveal latent similarities between tinnitus representations, and subsequently cluster those into distinct categories. Expected outcome: Unsupervised machine learning will identify distinct categories of tinnitus, which will allow us to further improve the efficiency and accuracy of uncovering tinnitus representations through improved stimulus generation and representation classification. \\

Our expected outcomes are (a) a substantially more detailed understanding of tinnitus experience and its variation across patients and (b) an novel, simple and efficient method for characterizing tinnitus experience in new patients. This will have a positive impact on individualized assessment of tinnitus and habituation therapies, and ultimately holds potential for improving health outcomes for tinnitus patients.

@@

\subsection*{C.1. Aim \#1: Uncover and reconstruct cognitive representations of tinnitus}

@@

\subsubsection*{Method}

@@

Classification of tinnitus into subtypes allows clinicians
and researchers to better understand the nature of the condition
by performing statistical comparisons between discrete, meaningful patient categories.
Surveys of tinnitus patients have identified sixteen qualitative subtypes of tinnitus
\cite{vajsakovicPrinciplesMethodsPsychoacoustic2021,meikleTinnitusArchiveArchive2004,stoufferCharacterizationTinnitusTinnitus1990},
but the utility of these categories Is limited by the fuzziness of linguistic categories and the musical vocabulary of the tinnitus patients.
Tinnitus subtypes are used to better explain and quantify the condition,
however qualitative subtypes are ill-defined.
While one tinnitus patient we interviewed, who has a degree in music performance,
was able to identify her tinnitus as sounding like a particular note played on the crotales,
a niche percussion instrument,
most tinnitus patients lack the depth of vocabulary to explain their tinnitus percept that precisely.
Accordingly, linguistic description of tinnitus can be vague and overlapping.
For instance, two of the subtypes in \cite{vajsakovicPrinciplesMethodsPsychoacoustic2021} are ``high tension wire'' and ``transformer noise.''
Is ``high tension wire'' meant in the colloquial sense of high-voltage power lines,
or does it mean the reverberation of a tightly-pulled string?
If the latter, then how is the sound of a high-voltage power line
disambiguated from the sound of a distribution transformer,
which is
mounted on a utility pole and connected directly to the high-voltage wires?
Instead of linguistic categories, we propose to use latent variable analysis
to reveal data-driven subtypes of tinnitus.
This will result in a new data-driven framework for categorizing and communicating
similarities and differences between tinnitus experiences.

@@

\begin{figure}[h] % Figure at bottom of the page ([b] argument, could be "t" for top or "h" for here)
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/variational-autoencoder.png}
	\caption{Diagram of a variational autoencoder (VAE) learning a representation of the digit ``2''.
	VAEs assume that the data, $x$, are generated by a directed graphical model, $p_\theta(x|z)$,
	called the decoder,
	and that the encoder, which maps from the input to the latent space, is learning an approximation $q_\phi(z|x)$. $\phi$ and $\theta$
	denote the parameters of encoder and decoder respectively.
	The latent representation, $z$, is interpreted as originating from a multivariate Gaussian distribution.
	By sampling from the latent distribution and passing it through the decoder, new data can be generated. Figure from \cite{IntroductionAutoencoders2020}.}
	\label{fig:vae}
\end{figure}

@@

While we anticipate remarkable performance gains in the number of samples required for accurate representation
through compressive sensing,
it may be clinically useful to identify only the subtype of tinnitus
(\ie which cluster on the lower-dimensional manifold the representation belongs to)
to some convergent accuracy.
If only the subtype is needed,
we believe the number of samples required for convergent accuracy can be further reduced.
We will devise an algorithm to perform the data-driven tinnitus subtyping ``online''
while the subject performs the task.
The algorithm monitors goodness-of-fit into the subtype space by posterior probability
and concludes the task when convergence is reached.
Through this process, both patient and clinician time is used efficiently, while still providing high-fidelity quantitative medical data
for use in diagnosis and treatment.

@@

\section*{D. Involvement of Undergraduate Students and Other Trainees}

The proposed projects will engage undergraduate students and other trainees extensively throughout the study period.
In 2018, Worcester Polytechnic Institute (WPI) was ranked by the Wall Street Journal as \#1 in Universities that best blend research and teaching.
WPI has several mechanisms that facilitate integrating students into research, and I have a longstanding personal interest and history of doing so.
Additional details about the numbers of students who will participate in aspects of this research, and their specific roles, are found in the Personnel Justification.
Briefly, students will contribute to the proposed research through three mechanisms: 
(1) class projects, (2) as individual researchers in my lab, and (3) through Major Qualifying Projects (MQPs).
I will personally oversee undergraduate student involvement, along with my Doctoral student Alec Hoyland.

\paragraph{Class Projects}
Students studying biomedical engineering must complete a rigorous data analysis (BME2211) course in their sophomore year. As part of that course, students complete a term-long project involving the collection of novel data sets and application of statistical modeling techniques to those data. All aims described here provide an ideal opportunity to engage students in the application of key methods in the proposed project. Students from this class will pilot the yes-no response protocols using existing code, and subsequently apply reverse correlation to the acquired data. Approximately 125 students per year take BME 2211.

\paragraph{Individual Researchers}
Each year my laboratory hosts 3-6 undergraduate student researchers who contribute to ongoing projects in the lab. These students typically start by volunteering 5-10 hours/week. However, many of them complete independent study projects and work (paid) for me over the summer, both of which require a high level of commitment and interest. Students will work on several aspects of the proposed project. Each will be given a specific experimental or technical question to answer that represents a subset of one of the Aims. For example, in Aim 2, students could be tasked to examine the effect of algorithmic parameters on compressive sensing reconstruction efficiency, in the interest of finding ideal parameter settings for reconstructing tinnitus representations on our collected data. Another pair might be tasked with determining the number of distinct tinnitus subtypes implied by the distribution of data.

Our department hosts an NSF-sponsored Research Experiences for Undergraduates (REU) program, which includes support for a female or underrepresented minority student from another University. One REU student per year will work on a project related to this research. The REU program includes structured mentoring (pairing undergraduate with graduate students, completion of a mentoring agreement, mentor training, etc.) and professional development activities. All summer undergraduate students hosted by my laboratory will be able to participate in these REU-affiliated programs, whether they are formally part of the REU or not.

\paragraph{Major Qualifying Projects (MQPs)}
WPI is a nationally recognized leader in project-based learning at the undergraduate level. Our curriculum is built around open-ended, team-based projects that must be completed by every student as a graduation requirement. During the senior year, students must complete their MQP, a project that is equivalent to 9 semester credit-hours. MQPs are performed in teams of 3-5 students.
I typically mentor multidisciplinary teams and students pursuing double majors in biomedical engineering and computer science. Each year as part of my regular teaching load, I mentor 4-12 MQP students (1-2 teams), meeting weekly for the entire academic year. Over the past 2 years at WPI, I have mentored 4 MQP teams (>15 students). During the award period, I expect to mentor two MQP teams each year that pursue projects that directly contribute to the Aims of this study or address questions related to the study goal.

\paragraph{Other Mechanisms of Student Engagement}
In addition to these formal mechanisms to engage students in research, students will have an opportunity for clinical exposure through our collaboration with Massachusetts Eye and Ear Infirmary.

MEEI is accessible from the WPI campus via commuter rail. Our collaborative investigator commonly interacts with tinnitus patients, and patients with auditory impairments of many different types. Students who participate in projects with Dr. Troy will each spend time at MEEI, and will be encouraged to attend seminar talks at MEEI. We believe this exposure to both the engineering and clinical aspects of tinnitus issues will provide students with a rich educational experience that will encourage them to pursue research in the future.
